Approximately Exact Calculations for Linear Mixed Models

### 2015-05-20 ms.tex
incidence of multimodality

It is known (e.g.\ \citealt{henn&hodges:2014}) that $\logRLss$\  
can have multiple maxima and that existing general purpose algorithms for linear mixed models may fail to find all of them, as shown by examples in \cite{hodges:2013}, \cite{henn&hodges:2014}, and elsewhere.
\cite{mullen:2014} examines 18 optimization functions available in \R; tests them on 48 objective functions, admittedly more complicated than $\logRLss$; and finds that even the best of them fail in over 10\% of the cases.  \cite{henn&hodges:2014} examine conditions under which multiple maxima occur in posterior densities and conclude ``\dots second maxima in posterior distributions therefore may be more common than reports in the literature would suggest."  Thus, failure to find local and global maxima may be common.

>>>>

%It is known (e.g.\ \citealt{henn&hodges:2014}) that $\logRLss$\  
%can have multiple maxima and that existing general purpose algorithms for linear mixed models may fail to find all of them, as shown by examples in \cite{hodges:2013}, \cite{henn&hodges:2014}, and elsewhere.
%\cite{mullen:2014} examines 18 optimization functions available in \R; tests them on 48 objective functions, admittedly more complicated than $\logRLss$; and finds that even the best of them fail in over 10\% of the cases.  \cite{henn&hodges:2014} examine conditions under which multiple maxima occur in posterior densities and conclude ``\dots second maxima in posterior distributions therefore may be more common than reports in the literature would suggest."  Thus, failure to find local and global maxima may be common.  
It is known (e.g.\ \citealt{henn&hodges:2014}) that $\logRLss$
can have multiple maxima,
though the incidence of multiple maxima is unknown.  Multi-modal posterior distributions arise readily from conflict between the likelihood and prior;  \cite{Liu&Hodges:2003} explores an important such conflict in detail and \cite{wakefield:1998} gives a naturally-occurring example explored further in \cite{henn&hodges:2014}.  Multiple maxima in restricted likelihoods have received far less attention and any statement about their incidence would be speculation.  To our knowledge, two naturally-occurring cases have have been reported, in \cite{welham_thompson:2009} and \cite{reiss_etal:2014};  \cite{henn&hodges:2014} report an artificial case and give a recipe for manufacturing examples.

As for numerical optimizers, it is known that existing general purpose algorithms for linear mixed models may fail to find all of the local maxima of $\logRLss$, as shown by examples in \cite{hodges:2013}, \cite{henn&hodges:2014}, and elsewhere.  \cite{mullen:2014} examines 18 optimization functions available in \R, tests them on 48 objective functions (admittedly more complicated than $\logRLss$) and finds that even the best of them fail in over 10\% of the cases.  \cite{henn&hodges:2014} examine conditions under which multiple maxima occur in posterior densities and conclude ``\dots second maxima in posterior distributions therefore may be more common than reports in the literature would suggest."  Thus, failure to find local and global maxima may be common, though with available tools it is extremely laborious to determine whether multiple maxima are present and where they are.
